SPARQL on the Open, Decentralised Web

### 2016-07-22 dissertation/conclusions.tex
Helen's corrections

As noted in Section~\ref{sec:motivation} it is clear that evaluating
arbitrary SPARQL queries are in general unsustainable, and it is
likely that the problems in Section~\ref{sec:perfproblems} are partly
to blame for it. Our work, and most of the related work in
Section~\ref{sec:related} approach this work from the direction of
moving stress from the server, but it is also clear that another way
to address Problem~\ref{prob:sparqlcomplex} is for the SPARQL engine
to identify queries it would be too heavy to answer already when
parsing or planning, and also expose that in the service description
so that clients can avoid asking them. The HTTP/2 specification
\cite{rfc7540} has some generic facilities that may be helpful:

>>>>

As noted in Section~\ref{sec:motivation} it is clear that evaluating
arbitrary SPARQL queries are in general unsustainable, and it is
likely that the problems in Section~\ref{sec:perfproblems} are partly
to blame for it. Our work, and most of the related work in
Section~\ref{sec:related} approach this problem from the direction of
moving stress from the server, but it is also clear that another way
to address Problem~\ref{prob:sparqlcomplex} is for the SPARQL engine
to identify queries it would be too heavy to answer already when
parsing or planning, and also to expose that in the service description
so that clients can avoid asking them. The HTTP/2 specification
\cite{rfc7540} has some generic facilities that may be helpful:

### 2016-07-22 dissertation/introduction.tex
Helen's corrections

\begin{enumerate}
\item Introduction of a path to critical practice of evaluations, that makes
  use of contemporary statistical techniques, to establish a practice
  that can be used to refute assertions on performance.
\item A didactical experiment to help researchers understand the statistics.
\item The novel application of a well established method in
  statistics, rarely used in Computer Science, to SPARQL endpoint
  evaluation.

>>>>

\begin{enumerate}
\item Introduction of a path for critical practice of evaluations, that makes
  use of contemporary statistical techniques, to establish a practice
  that can be used to refute assertions on performance.
\item A didactical experiment to help researchers understand the statistics.
\item The novel application of a well-established method in
  statistics, rarely used in Computer Science, to SPARQL endpoint
  evaluation.

### 2016-07-22 dissertation/papersincontext.tex
Helen's corrections

Its main weakness is how it deals with authentication: The basic idea
is that when a client visits a resource, it will be told what
write operations it can perform. However, write operations will rarely
be permitted unless the client has authenticated and been authorised
to do so, and so, it assumes that credentials will be supplied with
the first read request. While the HTTP standard permits clients to
supply credentials with such requests, in practice, it rarely if ever
happens, possibly for privacy reasons. Thus, the server will have to
challenge the client first, but this should not happen on a resource
that the client is permitted to read. Thus, the proposed hypermedia
needs a property so that the client understands what to do if it wants
to do write operations.

@@

To understand the context of this study, I noted none of the
benchmarks that was in use for the evaluation of SPARQL engines could
be used for basic statistical hypothesis testing, and clearly,
hypothesis testing must be a fundamental requirement. Then, I noted
substantial criticism forwarded by
\cite{Duan:2011:AOC:1989323.1989340} and \cite{MontoyaVCRA12}.

However, this study merely scratches the surface. To establish a
Design of Experiment-based methodology would be a multi-year project
in its own right, clearly beyond the scope of this
dissertation. Unfortunately, this also implies that I do not have the
rigorous evaluation that I would have preferred to have in this
dissertation, it is unattainable.

@@

This work resulted in a workshop paper entitled ``Pushing complexity
down the stack'', presented in the Developers Workshop at the
International Semantic Web Conference, by the lead author Gregory Todd
Williams, we were the two authors. The reference is
\cite{williamspushing}.

This work was the outcome of several hackathons that I had organised
in the Perl community. It is not very well known that Gregory Todd
Williams' SPARQL implementation written in Perl is one of the SPARQL
1.1 reference implementations. It is also fully compliant with
the specification. Through our work, and the availability of
traits-based programming in Perl, see \cite{traits}, we had extensive
discussions on how to leave optimisations to lower levels, e.g. the
underlying database engine, when they had the potential to provide a
better plan or cost estimate. While it is certainly possible to do
this without this new programming paradigm, traits made the
methodology much more comprehensible, and so helped the subsequent
work substantially.

@@

Dissatisfaction with the methodology used in the literature,
specifically benchmarking, to test assertions of performance lead to
exploration of statistical methods referenced in
Section~\ref{sec:condoe}, and to organise the Empirical 2014 Workshop
at the Extended Semantic Web Conference (which I was unfortunately
prevented from attending myself). Still, frustrated by the slow
progress of the last 16 years, and difficulty with comparing methods
to the more mature natural sciences, I still felt the need to express
a philosophical view of science, which I did by submitting an article
to the ``Negative or inconclusive results in Semantic Web'' at the
Extended Semantic Web Conference 2015. The essay, see
\cite{kjernsmo_noise_2015} was tangential to the Call for Papers, but
the chairs decided to accept it as an interview. With Ruben Verborgh
in the role of the interviewer, we reformulated it as
\cite{kjernsmo_noise_2015_interview}, which was published in the
proceedings and is included in the dissertation. It was performed with
Jacco van Ossenbruggen as the interviewer in front of the
audience. The essay, at
\url{http://folk.uio.no/kjekje/2015/noise-essay.html} is however a
more thorough account.

>>>>

Its main weakness is how it deals with authentication: The basic idea
is that when a client visits a resource, it will be told what
write operations it can perform. However, write operations will rarely
be permitted unless the client has authenticated and been authorised
to do so, and so, it assumes that credentials will be supplied with
the first read request. While the HTTP standard permits clients to
supply credentials with such requests, in practice, it rarely if ever
happens, possibly for privacy reasons. Thus, the server will have to
challenge the client first, but this should not happen on a resource
that the client is permitted to read. Therefore, the proposed hypermedia
needs a property so that the client understands what to do if it wants
to do write operations.

@@

To understand the context of this study, I noted none of the
benchmarks that were in use for the evaluation of SPARQL engines could
be used for basic statistical hypothesis testing, and clearly,
hypothesis testing must be a fundamental requirement. Then, I noted
substantial criticism forwarded by
\cite{Duan:2011:AOC:1989323.1989340} and \cite{MontoyaVCRA12}.

However, this study merely scratches the surface. To establish a
Design of Experiment-based methodology would be a multi-year project
in its own right, clearly beyond the scope of this
dissertation. Unfortunately, this also implies that this dissertation
does not feature an evaluation that lives up to the rigorous standards
I am advocating as it is unattainable within the time frame of this
work.

@@

This work resulted in a workshop paper entitled ``Pushing complexity
down the stack'', presented in the Developers Workshop at the
International Semantic Web Conference, by the lead author Gregory Todd
Williams. The reference is \cite{williamspushing}.

This work was the outcome of several hackathons that I organised
in the Perl community. It is not very well known that Gregory Todd
Williams' SPARQL implementation written in Perl is one of the SPARQL
1.1 reference implementations. It is also fully compliant with
the specification. Through our work, and the availability of
traits-based programming in Perl, see \cite{traits}, we had extensive
discussions on how to leave optimisations to lower levels, e.g. the
underlying database engine, when they had the potential to provide a
better plan or cost estimate. While it is certainly possible to do
this without this new programming paradigm, traits made the
methodology much more comprehensible, and so helped the subsequent
work substantially.

@@

This direction was motivated by a dissatisfaction with the methodology
used in the literature, specifically benchmarking, to test assertions
of performance. That lead to exploration of statistical methods referenced
in Section~\ref{sec:condoe}, and to the organising of the Empirical 2014
Workshop at the Extended Semantic Web Conference (which I was
unfortunately prevented from attending myself). Still, frustrated by
the slow progress of the last 16 years, and the difficulty in comparing
methods to the more mature natural sciences, I still felt the need to
express a philosophical view of science, which I did by submitting an
article to the ``Negative or inconclusive results in Semantic Web''
workshop at the Extended Semantic Web Conference 2015. The essay, see
\cite{kjernsmo_noise_2015} was tangential to the Call for Papers, but
the chairs decided to accept it as an interview. With Ruben Verborgh
in the role of the interviewer, we reformulated it as
\cite{kjernsmo_noise_2015_interview}, which was published in the
proceedings and is included in the dissertation. It was performed with
Jacco van Ossenbruggen as the interviewer in front of the
audience. The essay, at
\url{http://folk.uio.no/kjekje/2015/noise-essay.html} is however a
more thorough account.
